### 2015-04-24 doc/thesis_puc/review.tex
Adding some terms to be defined in the glossary

Optional type systems \cite{bracha2004pluggable} are an approach for
plugging static typing in dynamically typed languages.
They use optional type annotations to perform compile-time type checking,
though they do not influence the original run-time semantics
of the language.
This means that the run-time semantics should still catch type errors
independently of the static type checking.
For instance, we can view the typed lambda calculus as an optional
type system for the untyped lambda calculus, because both have the
same semantic rules and the type system serves only for discarding
programs that may have undesired behaviors \cite{bracha2004pluggable}.

@@

Gradual typing is similar to the approaches proposed by
Abadi and associates \cite{abadi1989dts} and Thatte \cite{thatte1990qst}
by including a dynamic type in a statically typed language.
However, these three approaches differ in the way they handle the
dynamic type.
While Siek and Taha \cite{siek2006gradual} rely on the consistency relation,
Abadi and associates \cite{abadi1989dts} rely on type equality with explicit
projections and injections, and Thatte \cite{thatte1990qst} relies on subtyping.

@@

Figure \ref{fig:consistent-subtyping} combines the consistency and subtyping
relations to compose the consistent-subtyping relation for the simply typed
lambda calculus extended with the dynamic type $?$.
When we combine consistency and subtyping, we are making subtyping handle
what casts are safe among static types, and we are making consistency
handle the casts that involve the dynamic type $?$.
The consistent-subtyping relation is not transitive, and thus
the dynamic type $?$ is not neutral to this relation.

So far, gradual typing looks like a mere formalization to optional
type systems, as a gradual type system uses the consistency or
consistent-subtyping relation to statically check the interaction
between statically and dynamically typed code, without influencing
the run-time semantics.

@@

There are some projects that incorporate gradual typing into some
programming languages.
Reticulated Python \cite{reticulated,vitousek2014deg} is a research
project that evaluates the costs of gradual typing in Python.
Gradualtalk \cite{allende2013gts} is a gradually-typed Smalltalk
that introduces a new cast insertion strategy for gradually-typed
objects \cite{allende2013cis}.
Grace \cite{black2012grace,black2013sg} is a new object-oriented,
gradually-typed, educational language.
In Grace, modules are gradually-typed objects, that is, modules
may have types and methods as attributes, and can have a state
\cite{homer2013modules}.
ActionScript \cite{moock2007as3} is one the first languages that
incorporated gradual typing to its implementation and
Perl 6 \cite{tang2007pri} is also being designed with gradual typing,
though there is few documentation about the gradual type systems
of these languages.

>>>>

Optional type systems \cite{bracha2004pluggable} are an approach for
plugging static typing in dynamically typed languages.
They use optional type annotations to perform compile-time type checking,
though they do not affect the original run-time semantics
of the language.
This means that the run-time semantics should still catch type errors
independently of the static type checking.
For instance, we can view the typed lambda calculus as an optional
type system for the untyped lambda calculus, because both have the
same semantic rules and the type system serves only for discarding
programs that may have undesired behaviors \cite{bracha2004pluggable}.

@@

Gradual typing \cite{siek2006gradual} is similar to two previous
approaches \cite{abadi1989dts,thatte1990qst}, because they also include
a dynamic type in a statically typed language.
However, these three approaches differ in the way they handle the
dynamic type.
While Siek and Taha \cite{siek2006gradual} rely on the consistency relation,
Abadi and associates \cite{abadi1989dts} rely on type equality with explicit
projections plus injections, and Thatte \cite{thatte1990qst} relies on subtyping.

@@

Figure \ref{fig:consistent-subtyping} combines the consistency and subtyping
relations to compose the consistent-subtyping relation for the simply typed
lambda calculus extended with the dynamic type $?$.
When we combine consistency and subtyping, we are making subtyping handle
which casts are safe among static types, and we are making consistency
handle the casts that involve the dynamic type $?$.
The consistent-subtyping relation is not transitive, and thus
the dynamic type $?$ is not neutral to this relation.

So far, gradual typing looks like a mere formalization to optional
type systems, as a gradual type system uses the consistency or
consistent-subtyping relation to statically check the interaction
between statically and dynamically typed code, without affecting
the run-time semantics.

@@

There are some projects that incorporate gradual typing into some
programming languages.
Reticulated Python \cite{reticulated,vitousek2014deg} is a research
project that evaluates the costs of gradual typing in Python.
Gradualtalk \cite{allende2013gts} is a gradually-typed Smalltalk
that introduces a new cast insertion strategy for gradually-typed
objects \cite{allende2013cis}.
Grace \cite{black2012grace,black2013sg} is a new object-oriented,
gradually-typed, educational language.
In Grace, modules are gradually-typed objects, that is, modules
may have types with methods as attributes, and they can also have
a state \cite{homer2013modules}.
ActionScript \cite{moock2007as3} is one the first languages that
incorporated gradual typing to its implementation and
Perl 6 \cite{tang2007pri} is also being designed with gradual typing,
though there is few documentation about the gradual type systems
of these languages.
